Safe Transferable Regions

### 2015-11-18 paper/type-system.tex
Sec 4. done.

Well-formedness and typing rules of \fbname establish the conditions
under which a region type is considered well-formed, and an expression
is considered to have a certain region type, respectively.
Fig.~\ref{fig:fb-staticsem} contains an illustrative subset of such
rules\footnote{The full formal development can be found in the
supplement}. The rules refer to a context ($\A$), which is a tuple of:
\begin{itemize}
\item A set ($\rhoenv \in 2^\rgn$) of static identifiers of regions
that are estimated to be live,
\item A finite map ($\aenv \in \tyvar \mapsto \fgjN$) of type
variables to their bounds\footnote{A bound of a type variable
($\tyvar$) in FGJ~\cite{fgj} is the class ($\fgjN$) that type variable
was declared to extend.  The map $\tyvar \mapsto \fgjN$ is denoted
$\Delta$ in FGJ formalism}, and
\item A constraint formula ($\phicx$) that captures the outlives
constraints on regions in $\rhoenv$.
\end{itemize}
$\A$ also contains $\rhomap$, 
% a map ($\rhomap \in \rgn \mapsto
% \{\CLOSED,\OPEN,\XFERRED\}$) from unique region identifiers of
% transferable regions to their current state in the FSM from
% Fig.~\ref{fig:region-fsm}.
which is primarily an artifact to facilitate the type safety proof,
and can be ignored while type checking user-written programs
in \FB. The context for the expression typing judgment also includes:
\begin{itemize}
\item A type environment ($\env \in x \mapsto \tau$) that contains the
type bindings for variables in scope, and 
\item The static identifier ($\rgn^a$) of the allocation context for
the expression is being typechecked.  
\end{itemize}
Like the judgments in FGJ~\ref{fgj}, all the judgments defined by the
rules in Fig.~\ref{fig:fb-staticsem} are implicitly parameterized on a
class table ($CT \in cn \mapsto D$) that maps class names to their
definitions in \FB.
% \begin{figure}

@@

To help state the type safety theorem, we extend \FB's expressions
with $\invalidexn$, and define the syntactic class of values:
\begin{smathpar}
\begin{array}{lclcl}
v & \in & \mathtt{values} & \coloneqq & \C{new}\; \fbN(\bar{v}) \ALT
\lambdaexp{\ralloc}{\rhoalloc\rhobar\,|\,\phi}{\taubar \; \xbar}{e}\\
  &     & & & \C{new}\; \RgnZT{\rgn}(v)\\
\end{array}
\end{smathpar}
Note that for $\C{new}\; \RgnZT{\rgn}(v)$ to be considered a value,
$\rgn \neq \toprgn$. The semantics reduces a $\C{new}\;
\RgnZT{\toprgn}(e)$ expression in the user program to a runtime
$\C{new}\; \RgnZT{\rgn}(v)$ value that is tagged with a unique
identifier ($\rgn$) for this region. A binding is also added to
$\rhomap$, mapping $\rgn$ to the ``closed'' state. The supplement
defines a type rule for such values, allowing them to be typed.
Type safety theorem is now stated thus:
\begin{theorem}
\emph{(\textbf{Type Safety})}
\label{thm:fb-type-safety}
$\forall e, \tau, \rhoenv, \rhomap, \rgn$, such that $\rgn \in
\rhoenv$ and $\tywf{\rhoenv}{\phicx}$, if $\hastyp{\emptyA,\rgn,
\cdot}{e}{\tau}$, then either $e$ is a value, or $e$ raises an
exception ($\redstoo{(e,\rhomap)}{\invalidexn}$), or there exists an
$e'$ and a $\rhomap'$ such that $\redstoo{(e,\rhomap)}{(e',\rhomap')}$
and $\hastyp{\emptyASigp,\rgn, \cdot}{e}{\tau}$.
\end{theorem}

>>>>

Well-formedness and typing rules of \fbname establish the conditions
under which a region type is considered well-formed, and an expression
is considered to have a certain region type, respectively.
Fig.~\ref{fig:fb-staticsem} contains an illustrative subset of such
rules\footnote{The full formal development can be found in the
supplement}. The rules refer to a context ($\A$), which is a tuple of:
\begin{itemize}
\item A set ($\rhoenv \in 2^\rgn$) of static identifiers of regions
that are estimated to be live,
\item A finite map ($\aenv \in \tyvar \mapsto \fgjN$) of type
variables to their bounds\footnote{A bound of a type variable
($\tyvar$) in FGJ~\cite{fgj} is the class ($\fgjN$) that type variable
was declared to extend.  The map $\tyvar \mapsto \fgjN$ is denoted
$\Delta$ in FGJ formalism}, and
\item A constraint formula ($\phicx$) that captures the outlives
constraints on regions in $\rhoenv$.
\end{itemize}
$\A$ also contains $\rhomap$, 
% a map ($\rhomap \in \rgn \mapsto
% \{\CLOSED,\OPEN,\XFERRED\}$) from unique region identifiers of
% transferable regions to their current state in the FSM from
% Fig.~\ref{fig:region-fsm}.
which is primarily an artifact to facilitate the type safety proof,
and can be ignored while type checking user-written programs
in \FB. The context for the expression typing judgment also includes:
\begin{itemize}
\item A type environment ($\env \in x \mapsto \tau$) that contains the
type bindings for variables in scope, and 
\item The static identifier ($\rgn^a$) of the allocation context for
the expression is being typechecked.  
\end{itemize}
Like the judgments in FGJ~\cite{fgj}, all the judgments defined by the
rules in Fig.~\ref{fig:fb-staticsem} are implicitly parameterized on a
class table ($CT \in cn \mapsto D$) that maps class names to their
definitions in \FB.
% \begin{figure}

@@

To help state the type safety theorem, we define the syntactic class of values:
\begin{smathpar}
\begin{array}{lclcl}
v & \in & \mathtt{values} & \coloneqq & \C{new}\; \fbN(\bar{v}) \ALT
\lambdaexp{\ralloc}{\rhoalloc\rhobar\,|\,\phi}{\taubar \; \xbar}{e}\\
  &     & & & \C{new}\; \RgnZT{\rgn}(v)\\
\end{array}
\end{smathpar}
Note that for $\C{new}\; \RgnZT{\rgn}(v)$ to be considered a value,
$\rgn \neq \toprgn$. The semantics reduces a $\C{new}\;
\RgnZT{\toprgn}(e)$ expression in the user program to a runtime
$\C{new}\; \RgnZT{\rgn}(v)$ value that is tagged with a unique
identifier ($\rgn$) for this region. A binding is also added to
$\rhomap$, mapping $\rgn$ to the ``closed'' state. The supplement
defines a type rule for such values, allowing them to be typed.
Type safety theorem is now stated thus:
\begin{theorem}
\emph{(\textbf{Type Safety})}
\label{thm:fb-type-safety}
$\forall e, \tau, \rhoenv, \rhomap, \rgn$, such that $\rgn \in
\rhoenv$ and $\tywf{\rhoenv}{\phicx}$, if $\hastyp{\emptyA,\rgn,
\cdot}{e}{\tau}$, then either $e$ is a value, or $e$ raises an
exception ($\redstoo{(e,\rhomap)}{\invalidexn}$), or there exists an
$e'$ and a $\rhomap'$ such that $\redstoo{(e,\rhomap)}{(e',\rhomap')}$
and $\hastyp{\emptyASigp,\rgn, \cdot}{e}{\tau}$.
\end{theorem}
